Statistical Grand Slams

### 2023-12-14 Lahman Analysis/Formal Analysis/Draft1.qmd
Finished through conclusion

@@

Overall, our model for predict the HR production of MLB hitters appears to do a fair job. It does a good job at updating its predictions based on the data for each player. But, still keeps its predictions close to the mean unless there is strong evidence, like with Pat Tabler, to suggest that it should deviate. Furthermore, as we can see with the contrast between Robin Yount and Kendrys Morales, our model does not have to rely solely on what the priors tell it. Our model can update itself to either follow a typical career trajectory, as we saw with Robin Yount, and it can update itself to follow an atypical career trajectory, as we saw with Kendrys Morales. Thus, we can conclude that our model does a good job at fitting overall. However, this model is far from perfect. We could see with predictions from Pat Tabler that our model would not predict HR probabilities $\pi$ at a rate where gambling would be a wise endeavor.

>>>>

If modeling for the sake of learning was not reason enough to embark on this endeavor, there is the possibility of financial compensation also looming with this research question. If one is able to accurately predict the HR's that a player hits in a season, then they would probably be able to gain employment for a MLB team. HR's are the pinnacle of a hitter's trip to the plate, they are guaranteed to score at least one run and can lead to up to four being scored depending on how many people are on base. Thus, MLB teams which seek to maximize wins are likely looking to hire analysts who can accurately predict the HR's a player would hit so that they know who to target for drafting, signing, and trading. In addition, sports betting has recently become legal in many places in the United States. Thus, if someone is able to

@@

Overall, our model for predicting the HR production of MLB hitters appears to do a good job. It does a fine job at updating its predictions based on the data for each player. But, still keeps its predictions close to the mean unless there is strong evidence, like with Pat Tabler, to suggest that it should deviate. Furthermore, as we can see with the contrast between Robin Yount and Kendrys Morales, our model does not have to rely solely on what the priors tell it. Our model can update itself to either follow a typical career trajectory, as we saw with Robin Yount, and it can update itself to follow an atypical career trajectory, as we saw with Kendrys Morales. Thus, we can conclude that our model does a good job at fitting overall since it appears to take extreme careers in stride.

However, this model is far from perfect. We could see with predictions of $\pi$ for Pat Tabler that our model would not predict HR probabilities $\pi$ at a rate where gambling would be a wise endeavor. These estimates of $\pi$ for many players would be off by enough that the chances of a making money from a bet would be unlikely. But, gambling was never the ultimate motivation behind the prediction. This paper sought to improve on the methods of Fellingham and Fisher (2017), which we believe we did. While it would have been nice to have the computational resources to run this model for 200,000 iterations, we were able to get somewhat stable estimates for our parameters on $\frac{1}{20}$ of the iterations. That is, instead of needing 200,000 iterations to get stable estimates, we were able to get stable estimates using only 10,000 iterations. This suggests that our model is better Fellingham and Fisher's (2017) model and predicting HR's.

Other than having better computational power, this model could be improved with better data. We implicitly assume in this paper that the probability of an AB resulting in a HR remains constant throughout an AB. However, we know that this is not true. A variety of factors that happen during the AB can effect the probability that any given pitch, and thus AB, results in a HR. These factors include, but are not limited to, the type of pitch, the speed of the pitch, crowd noise, state of the game, etc. Having access to this data would be interesting because it would allow our model to update its predictions of a player hitting a HR both before and during an AB. Some of this data is available to the public, and incorporating it into future research is an example of an area where this model could be improved upon.

In the future, this model could easily be adapted to predict other outcomes too, such as hits, walks, strike outs, etc. Any batting outcome would lend itself nicely to this modeling structure. Thus, in the future, it would be interesting to see how this model could be applied to other areas of baseball for prediction.
